---
title: "Select variables that match a pattern"
author: "Siddharth"
date: "`r Sys.Date()`"
output: html_document
---

If your data contain lots of columns whose names share a similar structure, you can use partial matching by adding `starts_with()`, `ends_with()` or `contains()` in your `select()` statement.

```{r}

library(tidyverse)

install.packages("nycflights13")

library(nycflights13)
data(package="nycflights13")

```

```{r}

glimpse(weather)
glimpse(flights)

```

These selection helpers match variables according to a given pattern.

`starts_with()`: Starts with an exact prefix.

`ends_with()`: Ends with an exact suffix.

`contains()`: Contains a literal string.

`matches()`: Matches a regular expression.

`num_range()`: Matches a numerical range like x01, x02, x03.

`starts_with()` selects all variables matching a prefix and `ends_with()` matches a suffix:

```{r}

flights %>% select(starts_with("dep"))

flights %>% select(starts_with("dep")|starts_with("air") )

flights %>% select(ends_with("time"))

```

You can supply multiple prefixes or suffixes. Note how the order of variables depends on the order of the suffixes and prefixes:

```{r}

flights %>% select(starts_with(c("dep","sched")))
flights %>% select(ends_with(c("time","hour")))

```

`contains()` selects columns whose names contain a word:

```{r}

glimpse(planes)
planes %>% select(contains("en"))

glimpse(weather)
weather %>% select(contains("_"))

```

`starts_with(), ends_with(),` and `contains()` do not use regular expressions. To select with a regexp use `matches()`:

```{r}

weather %>% select(matches("[wi]"))

weather %>% select(matches("^w"))

```

To select a range, use `num_range()`.

```{r}

glimpse(billboard)
?billboard

billboard %>% select(artist,track,num_range("wk",1:5))

```

Selection helpers also work together with the `& operator` signifies the intersection and `| operator` signifies the union of two selections.

```{r}

flights %>% select(contains("_") & ends_with("time"))

flights %>% select(starts_with("dep") | ends_with("time"))


```
