---
title: "dplyr filter() function to select the observations"
author: "Siddharth"
date: "`r Sys.Date()`"
output: html_document
---

Load the library

```{r}

library(tidyverse)
library(nycflights13)

```

Check first few observations

```{r}

glimpse(flights)

```

select all flights on Jan 1st

```{r}

flights %>%
  filter(month==1,day==1)

```

R provides standard suit of Comparison operators:

-   `>` greater than

-   `>=` greater than equal to

-   `<` less than

-   `<=` less than equal to

-   `!=` not equal

-   *`==` equal*


Find all the flights that departed in November or December

    ```{r}

    flights %>% 
      filter(month==11 | month==12) %>%
      count(month)

# The above expression can also be written using %in% operator. `x %in% y` will 
# select every row where x is one of the values in y    
    
flights %>%
  filter(month %in% c(11,12))%>%
  count(month)
    
    ```
    
Arrived more than 2 hours late, but did not leave late

```{r}

flights %>%
  filter(arr_delay>120,!(dep_delay>0))

```

filter function excludes FALSE or missing values
```{r}

df <- tibble(x=c(1,NA,3))

df %>%
  filter(is.na(x)|x>1)

```


how many flights have missing dep_time
```{r}
sum(is.na(flights))

flights %>%
  filter(is.na(dep_time))%>%
  count()

```
