---
title: "Pipe operator"
author: "Siddharth"
date: "`r Sys.Date()`"
output: html_document
---

```{r }

library(tidyverse)

```

# The pipe operator - %>%
R is function programming language, so quite often you will end up in writing the nested function calls to create complex data frames. In order to address this concern, tidyverse has a very handy R Package, called magrittr, which lets you transform the complex nested function calls, into pipeline of operations that is easier to understand and to write. 

The operators pipe their left-hand side values forward into expressions that appear on the right-hand side, i.e. one can replace f(x) with x %>% f(), where %>% is the (main) pipe-operator. 

Basic piping:

 x %>% f is equivalent to f(x)
 x %>% f(y) is equivalent to f(x, y)
 x %>% f %>% g %>% h is equivalent to h(g(f(x))). In this example, three operations are performed to arrive at the desired result, and they are written in the natural order, same as order of execution.

# load mtcars data set
```{r}

head(mtcars)

```

Perform following operations:
 1. load mtcars data set, and group it by number of cylinders
 2. summarise the group by mean of Miles per gallon variable
 3. display the result in the descending order of the # of cylinders


# Option 1: Store each step in the process sequentially:

```{r pressure, echo=FALSE}
result <- group_by(mtcars, cyl)
result <- summarise(result, meanMPG = mean(mpg))
result <- arrange(result,desc(cyl))
result
```

# Option 2: Nesting the functions:

```{r}

result <- arrange(summarise(group_by(mtcars, cyl),meanMPG = mean(mpg)),desc(cyl))

result
```

# Option 3: use pipe operator:

```{r}

result <- mtcars %>% 
    group_by(cyl) %>% 
    summarise(meanMPG = mean(mpg)) %>%
    arrange(desc(cyl))

result

```

If we read %>% as “then”, the code from the previous section is very easy to digest as a set of instructions in plain English:

To get our result, take the mtcars data-frame, THEN
Group its entries by number of cylinders, THEN
Compute the mean miles-per-gallon of each group, THEN
arrange the result in descending order



