---
output:
  html_document: default
  word_document: default
---
---
title: "R Notation"
author: "Siddharth"
date: "`r Sys.Date()`"
output:
  html_document: default
 ---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load the Data Frame

To demonstrate the R notation, we will use *deck.csv* file, which contains 52 observations representing full deck of cards. Let's load the data set:

```{r}
getwd()
setwd("E:/Training Modules/introduction to R/R Notation")
getwd()
card_deck <- read.table("deck.csv",sep=',',header=TRUE, stringsAsFactors = FALSE)
```

### Check first few rows of data set

```{r}
head(card_deck)
```

## Positive Integers

R treats positive integers just as *ij* notation in linear algebra: `r`deck[i,j]\`, will return the value of the deck that is in *ith* row and *jth* column

```{r}

card_deck[1,1]

```

To extract more than one value use vector of positive integers

```{r}

card_deck[1,c(1,2,3)]

```

R's notation system is not limited to data frames. You can use the same syntax to select values in any R object.

```{r}

vec <- c(1,2,3,4,5,6)
vec[1:3]

```

### Negative Integers

R will return every index except the elements in the negative index.

```{r}

card_deck[-(2:52),  1:3]

```

R will return an error if you try to pair negative integer with positive integer in *same* index

```{r}
card_deck[c(1,-1),1]
```

### Zero

What would happen if you used zero as an index?

```{r}

card_deck[0,0]
```

### Blank Spaces

Useful for extracting entire row or column from a data frame

```{r}

card_deck[1,]
```

### Logical Values

If you use a vector of TRUE and FALSE as index, R will match each TRUE and FALSE to a row/column in data frame

```{r}

card_deck[1, c(TRUE, TRUE, FALSE)]

card_deck[c(FALSE,FALSE,TRUE),1:2]

```

### Names

If the columns have names, then you can extract the columns of data frame using its name

```{r}
card_deck[ , "suit"]

```


### Dollar Sign and Double Brackets

\$ symbol is used to extract values from data frame and lists. To select column from data frame, write the data frame's name and column name separated by \$.

```{r}
card_deck$value
```

```{r}
mean(card_deck$value)
median(card_deck$value)
```

sub-setting a list, \$ notation, single bracket and double brakcet

```{r}
lst <- list(number=c(1,2),logical=c(TRUE,FALSE),strings=c("a","b","c"))
lst

```

## Sub-setting list - Using single bracket.

for single bracket sub-setting, the result is smaller list within a list. This can cause problem, because many R functions do not work with the list. For. e.g. `sum(list[1])` will throw an error

```{r}
lst[1]
sum(list[1])
```

## \$ notation in the list

The above issue can be addressed by using \$ notation. When \$ notation is used, R will return selected values, with no list structure attached around them:

```{r}
lst$number
sum(lst$number)
```

## Sub-setting list - Using double bracket.

If the elements in the list do not have names, you can use two brackets to subset the list. This is similar to using dollar notation for sub-setting the list.

```{r}
lst[[1]]
sum(lst[[1]])
```

# Atomic vector used in example
```{r}
myvec <- c(1,4,56,21,14)
myvec

myvec[-2]
```

# Data frame used in example
```{r}
my_dataframe <- data.frame(person=c("Siddharth","Aarya","Meghana","Omkar","Atharva"),
age=c(42,40,17,14,1),
sex=factor(c("M","F","F","M","M")))

my_dataframe

my_dataframe[1,c(TRUE,TRUE,FALSE)]
```

# List used in example
```{r}
my_list <- list(my_matrix=matrix(data=1:4,nrow=2,ncol=2),logical=c(T,F,T,T),strings="hello",numbers=c(10:14))
my_list

my_list[1]
```
